---
title: "carib_results"
author: "Lennon Thomas"
date: "2/8/2018"
output: html_document
---

```{r setup, warning=FALSE , echo=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = F, warning = F, fig.width = 6.5)
```



```{r echo = FALSE}
# Load packages and functions-----------------------------------------------------------
rm(list = ls())

library(raster)
library(rgdal)
library(tmap)
library(ncdf4)
library(stringr)
library(broom)
library(fasterize)
library(sf)
library(tidyverse)
library(parallel)
library(R.utils)
library(readr)
library(pander)
#library(plotly)
library(plyr)

# Run settings -------------------------------------------------------------

## Set User (lennon/tyler)
user <- 'lennon'

if(user == 'lennon') { boxdir <- '/Users/lennonthomas/Box Sync/Waitt Institute/Blue Halo 2016/Carib_aqua_16/'}
if(user == 'tyler')  { boxdir <-  '../../Box Sync/Carib_aqua_16/'}

run_name = 'est_feb_8'  
# Load EEZ shapefile for plotting

EEZ = readOGR(dsn=paste(boxdir,"Suitability/tmp",sep = ""),layer="carib_eez_shape")

# Load suitablity results
s_areas<-gzfile(paste0(boxdir,"/results/Suitability/suitable_areas.rds"))

suit_areas<-readRDS(s_areas)

suit_df<-read_csv(paste0(boxdir,"/results/Suitability/suitable_area_df.csv"),
                  col_types = "iiddiic") %>%
  select(c("cell_no","study_area_km","suit_area_km","suit_index","eez","country"))

# Load sim data

data_folder <- paste0(boxdir,'results/',run_name,"/Data/")

data.names <- list.files(path = data_folder, pattern = ".nc")

data_files <- lapply(paste0(data_folder,"/",data.names),brick)

avg_growth<-data_files[[1]]

harv_cycle_length<-data_files[[2]]

no_cycles<-data_files[[3]]

stocking_n<-data_files[[4]]

rm(data_files)

# Load run results 
  #"calc_0.02" #run name reflects intital stocking density (calculated or fixed)and feed rates (as % body weight)

result_folder <- paste0(boxdir,'results/',run_name,"/Results")

file.names <- list.files(path = result_folder, pattern = ".csv")

result_files <- lapply(paste0(result_folder,"/",file.names),read_csv)

carib_supply<-result_files[[1]]

eez_supply_df<-result_files[[2]]

npv_df<-result_files[[4]]

supply_summary<-result_files[[6]]

rm(result_files)
```


## Suitability Analysis

We estimated `r sum(suit_df$suit_area_km,na.rm = TRUE)` km^2^, or `r round((sum(suit_df$suit_area_km,na.rm = TRUE))/(sum(suit_df$study_area_km,na.rm = TRUE)) *100,2)`% of the total study area (Fig.1).
```{r}
EEZ$MRGID<-{as.numeric(levels(EEZ$MRGID))[EEZ$MRGID]}

country_code<-as.data.frame(EEZ@data[c("MRGID","Territory1"),])

country<-raster(paste0(boxdir,"Suitability/tmp/carib_eez_raster.tif"))
  #rasterize(EEZ,suit_areas,field = 'MRGID')

suit<-as.data.frame(rasterToPoints(suit_areas)) %>%
  set_names(c("Longitude","Latitude","suitable"))

country_area<-area(suit_areas,na.rm=TRUE)

country_suitable<-zonal(country_area,country,fun='sum')

country_suitable<-as.data.frame(country_suitable)

country_s_area<-cbind(country_suitable,EEZ$MRGID,as.character(EEZ$Territory1)) %>%
 select(c(2,3,4)) %>%
 purrr::set_names(c("Area","MRGID","EEZ")) %>%
 arrange(desc(Area)) %>%
 select("EEZ","Area","MRGID")
```

`r as.character(country_s_area$EEZ[1])` had the largest suitable area and `r as.character(country_s_area[country_s_area$Area==0,1])` did not have any areas identified as suitable for offshore mariculture.


```{r Fig. 1 echo = FALSE}

EEZ@data$suitable<-country_suitable$sum
EEZ@data$mytext<-paste0("EEZ =",EEZ@data$Territory1, "/n", "Suitable area (km^2) = ", EEZ@data$suitable)

EEZ@data$id<-rownames(EEZ@data)
EEZ.points<-fortify(EEZ, region="id")
EEZ.df<-join(EEZ.points,EEZ@data,by="id")
EEZ.df$suitable<-ifelse(EEZ.df$hole==TRUE,NA,EEZ.df$suitable)
 
EEZ.df<-EEZ.df[EEZ.df$hole==FALSE,]

p<-ggplot(EEZ,aes(text = mytext))+ 
  geom_polygon( aes(long, lat, group = group, fill = hole), colour = "black", size = 0.1) + 
  scale_fill_manual(values = (c("lightblue", "white")),guide = FALSE) + 
  theme(legend.position="none") +
  theme_minimal() +
   geom_raster(data=suit,aes(x=Longitude,y=Latitude,fill=suitable),fill="darkred") +
    xlab("Longitude") +
  ylab("Latitude") +
  coord_fixed(xlim =c(-85.5,-57.4),ylim = c(9.95,30))

ggplotly(p)
  
##add ability to interactivy zoom into regions/
```


## Production

```{r}
npv_df$cell<-as.factor(npv_df$cell)
npv_df$eez<-as.factor(npv_df$eez)

eez_summary<- npv_df %>%
  filter(discounts==0,prices==9) %>%
  group_by(eez) %>%
 dplyr:: summarise(total_harvest = sum(harvest),
                   npv = sum(npv)) %>%
  ungroup()

EEZ<-merge(EEZ,eez_summary,by.x = 'MRGID',by.y = 'eez')


EEZ@data$id = rownames(EEZ@data)
  EEZ.points = fortify(EEZ, region="id")
  EEZ.df = join(EEZ.points, EEZ@data, by="id")
EEZ.df$total_harvest<-ifelse(EEZ.df$hole==TRUE,NA,EEZ.df$total_harvest)
EEZ.df$npv<-ifelse(EEZ.df$hole==TRUE,NA,EEZ.df$npv)

ggplot(EEZ.df)+ 
    geom_polygon(aes(x=long,y=lat,group=group,fill=total_harvest),col="black",lwd=0.5) +
  #geom_polygon(data = EEZ.df, aes(long, lat, group = group),fill =NA, colour = "black", size = 0.1) + 
  #scale_fill_manual(values=c("lightblue", "white"), guide=FALSE)+
  theme(legend.position="none",legend.title=element_text("Total 10 yr production (kg)")) +
 theme_minimal() +
    xlab("Longitude") +
  ylab("Latitude") +
  coord_fixed(xlim =c(-85.5,-57.4),ylim = c(9.95,30)) +
  scale_fill_continuous(na.value="white")

ggplot(EEZ.df)+ 
    geom_polygon(aes(x=long,y=lat,group=group,fill=npv),col="black",lwd=0.5) +
  #geom_polygon(data = EEZ.df, aes(long, lat, group = group),fill =NA, colour = "black", size = 0.1) + 
  #scale_fill_manual(values=c("lightblue", "white"), guide=FALSE)+
  theme(legend.position="none",legend.title=element_text("Total 10 yr production (kg)")) +
 theme_minimal() +
    xlab("Longitude") +
  ylab("Latitude") +
  coord_fixed(xlim =c(-85.5,-57.4),ylim = c(9.95,30)) +
  scale_fill_continuous(guide="colorbar",na.value="white",low="darkred",high="darkblue")

```

```{r}

bp_data<-npv_df %>%
  filter(prices==9)
bp_data$discounts<-as.factor(bp_data$discounts)

ggplot(bp_data,aes(x=discounts,y=npv))+
  geom_boxplot() +
  theme_minimal()+
  facet_grid(~eez) +
  xlab("Discount Rate") +
  ylab("10 yr NPV")

```

