library(knitr)
library(pandoc)

---
title: "Cobia Aquaculutre in the Caribbean"
output:
  html_notebook:
    toc: yes
  word_document:
    toc: yes
---


## Introduction

I) Summary and importance of global aquaculture production  

    a. Brief summary of ecological, social and economic benefits of aquaculture  
    
    b. Projected increase in seafood demand and aquaculture as the most likely means to fill this demand (i.e., limited ability of wild fisheries to increase production)  
    
    c. Description of factors that must be considered when determining aquaculture production potential of a given region (physical, ecological, social, and economical (?))~ To determine the sustainable aquaculture production capacity of an area or region, four dimensions must be considered:  ecological, social, physical, and production (McKindsey et al. 2006). The ecological capacity is defined as the maximum aquaculture production that can be sustained in a region without impacting ecological processes, services, populations or communities in an environment (Dunn et al. 2013). The social capacity of an area for aquaculture production is broadly defined as the maximum aquaculture production that can be developed in a region without adverse social impacts (REF). The physical capacity of a region relates to the amount of space available that meet the technological constraints of the aquaculture, and/or environmental requirements of the aquaculture infrastructure and species being cultured (Dunn et al.2013). The production capacity is typically defined at the farm level, and refers to the maximum production that is feasible given the aquaculture systems and technology available, species being cultured, and economic conditions.  
  
II)	The Caribbean as a region where aquaculture development could be particularly beneficial   
    
    a. The Caribbean Sea is home to 9% of the world’s coral reefs, and has a population of over 42 million people, many who depend on the area’s marine resources for subsistence and income (WRI 2006)  
    
    b. Caribbean seafood consumption and production (include information on seafood exports and imports)  
    
    c. Describe current aquaculture production occurring in the Caribbean  
    
    d. Recently, further development of aquaculture has been identified as a priority in the Caribbean  
    
    e. Potential for land-based and coastal aquaculture in the Caribbean is extremely limited (explain why…..limited space on land, limited freshwater, limited energy resources, environmental concerns with coastal aquaculture), suggesting the need to look to offshore aquaculture  
  
III) Offshore aquaculture is an emerging approach to mariculture where farms are some distance offshore  

    a. Overview of offshore aquaculture worldwide  
    
    b. Brief discussion of advantages of offshore aquaculture (specifically in the Caribbean)  
    
    c. Likely species/technology in the Caribbean  
    
        i.	Overview of cobia biology and mariculture  
        ii.	SeaStation submerged cages  
  
  
IV) Introducing our study  

    a. Offshore aquaculture is nearly untapped in the Caribbean, raising questions about what the actual potential is, how that potential varies across the many EEZs of the Caribbean, and what factors are currently most likely to be limiting development 
    
    b. Using high resolution spatial data, we develop a spatial bioeconomic model to estimate the total production potential (in terms of weight and profit) of offshore aquaculture in the Caribbean. We examine the spatial distribution of offshore aquaculture potential to identify ‘hotspots’ for potential future offshore aquaculture development, and how it is distributed across space. Additionally, we examine the variation of individual farm profitability between and across EEZ to determine the importance of site selection in aquaculture development and to quantify the potential benefits of strategic spatial planning.  We explore these results under several regionally scaled production capacity scenarios, and identify the main parameters constraining production in our model. 


## Methods  

### Description of study region

The area of focus for this study is the territorial waters surrounding the 28 island countries that comprise the Greater and Lesser Antilles of the Caribbean Sea (Figure 1). This area was defined using a shape file of the Exclusive Economic Zone (EEZ) boundaries for all island countries  from VLIZ (2014). 

```{r Caribbean EEZ tmap, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="**Figure 1.** The area included in this study is indicated in Blue. EEZ boundary data from VLIZ (2014)",fig.width=6.5, fig.height=4}

library(tmap)
library(rgdal)
data(World)

land = readOGR( dsn="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/EEZ",layer="carib_eez_shape")
country<-land@data$GeoName[c(1,2,3,4,5,8,9,10,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,34,35,36,37,40,41,42,43,44,45,48)]
land$la<-ifelse(land@data$GeoName %in% country,"lightblue","white")
 tm_shape(land)+
  tm_fill("la")+
  tm_borders()+
   tm_legend()
```


### Suitability Assessment

The first step in our analysis was to conduct an assessment to identify areas that were suitable for the development of offshore  mariculture. To determine what areas were suitable, we considered 6 factors (Table 1). To develop mariculture sustainably, conflicts between other users and uses of the marine environment should be minimized (Longdill et al. 2008). Additionally, offshore mariculture is only feasible in areas that meet the techinical requirements of the current offshore cage technology, without creating large additional installation and maitenance costs.  SeaStation is  designed for offshore aquaculture. Seastation, a submersible, self-tensioned, single rim cage, is the most widely used offshore aquaculture cage in the U.S. (Loverich 2000; Open Blue 2016).The minimum site depth for installation of a SeaStation cage, listed by the original manufacturer, is 25 m (OceanSpar 2013). The maximum suitable depth for cage installation for this analysis was set at 90 m.  In waters deeper than 90 m, cage installation and inspection of mooring and anchoring systems is more difficult and costly (Scott and Muir 2006). Spatial bathymetry data were obtained from Becker *et al.* (2009). SeaStation cages are suitable for areas experiencing currents velocities up to 1 $\frac{m}s$ (Loverich 2007).


To identify areas currently designated or utilized for other purposes, and unsuitable for offshore mariculture development, we obtained spatial data on: shipping activity, existing benthic oil structures, areas permitted for deep sea bed mining, and areas that have been designated as Marine Protected or Conservation Priority areas (Table 1). Areas that fell in the 10% of the highest relative shipping activity, designated for conservation or deep sea bed mining purposes, and with existing benthic oil structures were considered unsuitable for offshore mariculture development. To identify areas that met the technical requirements for the installation of offshore SeaStation cages we considered depth and current speed (Kapetsky et al. 2013). Bathymetry data were obtained from Becker *et al* 2009, and areas that fell within a 25-90 m depth were scored as suitable for offshore mariculture development. To identify areas with potential of high current velocities, e obtained 10 years (2005-2015) of monthly average current data from Ocean Surface Current Analysis Real-time (OSCAR) zonal and meridional surface current velocity estimates from JPL Physical Oceanography DAAC and developed by Earth Space Research (ESR 2009) (Table 1). The absolute maximum zonal and meridional velocities for each cell in our study area were calculated, and any cell with a maximum zonal or meridional velocity greater than 1 $\frac{m}s$ were scored as unsuitable for the development of offshore mariculture


All spatial data layers listed in Table 1 were converted to raster format with a 1 km^2^ spatial resolution. Data files that had a resolution > 1 km^2^ were interpolated to the 1 km^2^ grid using 'nearest neighbor technique'. All 1 km^2 cells from each layer that fell outside of the suitable threshold, were given a score of zero. We then stacked all data layers in a boolean overay to identify areas suitable for offshore mariculture development based on the criteria in Table 1.   


```{r kable, echo=FALSE, warning=FALSE}
data<-read.csv("table_1.csv",check.names=FALSE)
#knitr::kable(head(data),caption="Table 1. Spatial data layers used in suitablility analysis to #identify cells suitable for cobia offshore mariculture",row.names=FALSE, #align=c('c','c','l','c'),padding=10)
library(pander)
set.alignment("center")
names(data) <- pandoc.strong.return(names(data))
pander::pandoc.table(data,justify=c('center','center','center','left','left','center'),split.cells=c(9,9,9,40,10,10),split.tables=Inf,caption=" Spatial data layers used in suitablility analysis to #identify cells suitable for cobia offshore mariculture")
```

### Farm Design
#### Individual Farm Design

SeaStation cages are typically configured using a grid mooring system that includes grid, anchor, and mooring lines secured at varying distances from the cages (Fredricksson et al. 2004; Xu et al. 2015).We developed a hypothetical farm design for a 1 km^2^ cell that has has 16 6,400 m^3^ SeaStation cages configured in two eight-cell grid mooring systems that occupies a total space of approximately (~0.48 km^2^) and has a total cage capacity of 102,400 m^3^ (Figure 2). The cages are held in position by the mooring system at depths of 15-20 m (Loverich 2010).

This cage configuration is similar to a 0.4 km^2^ offshore mariculture farm located off the coast of Kona that has an eight-cage SeaStation with a total cage volume of 64,000 m^3^ (Sims et al. 2013). The total cage volume per unit of total farm area of this hypothetical design falls within the range of total cage volume per unit area of total farm area offshore mariculture farm off the coast of Kona (Sims et al. 2013), and another offshore mariculture farm using SeaStation cages in the Gulf of Maine (Fredreicksson et al. 2004; DeCew et al. 2010). Additionally, the total space occupied by the farm's infrastructure (0.48 km^2^) follows the guidelines issued in NOAA's Fishery Management Plan (FMP) for offshore aquaculture development in the Gulf of Mexico that total farm area should be twice the size of the total area occupied by farm infrastructure (FMP 2009).

To maintain a steady supply of product, commercial aquaculture farms typically stage the stocking of cages so that indivdiuals of harvestable size are available year round. Cobia take on average, 12 months to grow from fingerling to a harvestable size ().A intial stocking density of 7.2 grams per m^3^ produced higher growth rates for cobia in the Caribbean than cages stocked at a higher initial density (Bennetti et al. 2010). We assume that the stocking of individual cages within a farm is staged at monthly time step. This means that for any given month, there will be between 1 and 5 cages with cobia at a harvestable age. We set the intial stock density of cages at 7.2 grams per m^3, which is equivalent to 2.3 juveniles with an average weight of 3 grasm (Bennetti et al. 2010).

### Bioeconomic Model

We applied a bioeconomic model to estimate the production (in terms of cobia biomass (mt) and profit ($)) of individual farms developed in each of the 1 km^2^ cells that were identified as suitable for the development of offshore aquaculture.  Only farms that generated profit over a ten year period were ultimately considered feasible for development.

We used a thermal performance curve (TPC) to model temperature dependent growth of cobia (Klinger et al. . Temperature is one of the primary abiotic factors controlling growth in fish, including cobia (Benetti et al. 2010), and something that cannot be controlled in offshore mariculture grow-out settings. We used 10 years of satellite-based sea surface temperatures (SST) (NASA 2014) to calculate 1 km2 resolution spatial data layers of average monthly SST. 

We assume a t_min_ of 22 C and a t_max_ of 32 C (REF). Monthly biomass $BM$ is then calculated for each month, at each individual farm as:

* INSERT DANE'S THERMAL PERFORMANCE CURVE EQUATION *

where $D$, is the initial stocking density ($\frac{kg}{m^3}$ ) of each farm, and $M$, the monthly natural mortality rate is assumed fixed at XX.
 
Monthly biomass is then summed, to determine total annual biomass ($B_t$) at each farm.

Where $B_t$ estimates biomass (mt) in month $t$ as a function of average monthly sea surface temperature, assuming an initial stocking density of XX, and a natural mortality ($M$) rate of XX (Bennetti et al. 2010). 

Using parameters we derived from the literature, the economic model estimates total annual operating costs ($TC_t$) for year 1 ($t = 1$) as:  
$TC_{t = 1} = SC + OC_{t = 1}$

Where $SC$ represents capital expenditures (start up costs) as:  

$SC_{t=1} = E_{depth} + I_{distance,EEZ} + P_{EEZ}$

where $E$ represents materials and equipment costs as a function of depth, I represent installation costs as a function of distance from shore and EEZ, and $P$ represents the cost of an aquaculture permit or lease as a function of EEZ (Table 2). 

$OC_t$ represents annual operating costs at each year $t$:

$OC_t = (u_{wave_exposure} * w * d_{EEZ}) + (\frac{distance}{b} * w * d_{EEZ} * r) + (distance * b * g_{EEZ} * e * t) + (f_t +h_t)$

Where $u$ is the hours required at each site as a function of wave exposure, $w$ is the number of workers required to service a farm, and $d$ is the hourly wage per worker as a function of EEZ, $distance$ is distance from shore (km), $b$ is average boat speed to site, $r$ is the number of trips required per month for running the farm, $g$ is the cost of fuel as function of EEZ, $e$ is the average fuel efficiency of the vessel, $f$ is the cost of fingerlings in year $t$, and $h$ is the cost of feed in year $t$. The global market for cobia is currently very small, with just 8.5% of production (4,555 MT) exported in 2012. We assume that cobia aquaculture in the Caribbean will be an export commodity and thus increased production in the Caribbean will affect the global price of cobia $c_t$ as follows:

$c_{t+1} = c_t - 0.8c_{t}(\frac{\Delta_{x}}{x_{t-1}})$

Where $c_{t+1}$ is the global price of cobia (\$) in the following year and $\frac{\Delta_{x}}{x_{t-1}}$ is the percent change in Caribbean production from the previous year. The constant 0.8 represents the price elasticity of demand and suggests that for every 1% increase in price there is a 0.8% decrease in consumption.

Total annual revenue, $TR$, for each year $t$,  for each farm was calculated as :

$TR_t = ( B_t c_t) – (B_t c_t j_{EEZ})$ 

Where $j$ is the export tax as a function of EEZ. 

The total profit ($TP$) for each farm over $t$ years was calculated as:

$TP = \displaystyle\sum_t^{10}TR_t -\biggl[\displaystyle\sum_t^{10}SC_t + \displaystyle\sum_t^{10}OC_t\biggr]$

The economic model uses parameters derived from the literature to estimate start-up costs and annual operational costs for each 1 $km^2$ farm, and estimates farm profits based on production (mt) estimated from the biological model, and an assumed fixed market price for cobia. (Table 2). Startup costs include: 

We calculated the total annual revenue and biomass (mt) generated by each farm over a ten year period, and only farms that generate positive revenue will be considered feasible. Feasible farms within each EEZ were then ranked according to revenue (highest to lowest). We then examined the range of estimated revenues generated by farms within an EEZ, and compared the highest and lowest revenue generating farms across EEZs.


### Production Capacity (regional ecological carrying capacity)

Determining the aquaculture production level that a given region (in this case, EEZ) can sustain without resulting in negative impacts to the surrounding ecosystem and environment is difficult, and will depend on the specific local oceanographic conditions and ecosystem dynamics. Although models have been developed to simulate conditions of a local area and to generate impacts of given aquaculture production scenarios on the surrounding ecosystem, currently, limited empirical evidence exists to estimate the ecological carrying capacity for aquaculture production at a regional scale due to lack of extensive offshore aquaculture development in comparable ecosystems. This production potential is often assessed at the farm level by taking in to account localized oceanographic and ecological characterstics at a specific aquaculture site location (REF). Far less is known about potential cumulative ecosystem impacts of aquaculture development at the regional scale.
```{r GOM FMP}
Totalarea<-35643
Totalproduction<-29000
Productionrate<-Totalproduction/Totalarea
```

Guidelines for aquaculture development in U.S. waters of the Gulf of Mexico listed in a recent Fishery Management Plan stated that all aquaculture farms should be a minimum of 3 km apart and set a precautionary production limit of `r prettyNum(Totalproduction)` mt for a total area of `r prettyNum(Totalarea)` km^2^ identified for aquaculture developement, or `r prettyNum(Productionrate)` mt/ km^2^. This production limit was identified as precautionary and based on historical wild fisheries capture for the region, due to lack of better methods or information to inform the ecological capacity of the region. To determine where this specified limit on production per unit area compared to production per unit area in other regions, we conduced a global production assessment to determine the range of aquaculture per unit of suitable area for all EEZs. We obtained total marine aquaculture production by country for 2014 from FAO and calculated the total area in a suitable depth range (0-90 m) for aquaculture by EEZ using global 1 km^2^ resolution bathymetry data from Becker *et al* 2009.

Because determining an ecological production capacity at a larger spatial scale (ie. EEZ) is difficult, we investigate a range of potential regional production limit scenarios at the EEZ scale for the Caribbean:

1. No production limit- There is no regionally specified production limit
2. High production limit- We set the production limit by identifying the EEZ with highest production per unit area globally from our  production assessment and multiply that value by the total area identfied as suitable for offshore aquaculture development in each EEZ.
3. Precautionary limit- The production limit for each EEZ is specified by multiplying the production per unit area value specified for the Gulf of Mexico (`rProductionrate) by the total area identified as suitable for offshore aquaculture development in  each EEZ.

Under each scenario, individual farms were developed in order of highest annual revenue. Once a single farm is developed, all farms falling within 3 km in any direction from that farm are excluded (not developed).

We then examined and compared the following under each production capacity scenario:
•	Total production (in terms of profit and weight) for the entire study (Caribbean) region
•	Total production (in terms of profit and weight) for each EEZ
•	Range of individual farm production within EEZs (to the see potential gains of strategic spatial planning) and across EEZs (to determine the island countries with the greatest potential for offshore aquaculture development)

Finally, we determined the main factors contributing to variability of farm and EEZ level aquaculture production by performing a suite of sensitivity analyses, in an effort to identify the current major hurdles to the development of offshore aquaculture in the Caribbean. 

	
## Results
