---
title: "Charting an Aquaculture Future for the Caribbean"
author: "Lennon R. Thomas, Tyler Clavelle, Dane Klinger, Rebecca Gentry, Jamie Afflerbach,
  and Sarah E.Lester"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
  word_document:
    toc: yes
bibliography: carib_aquaculture.bib
---


# Introduction

I) Summary and importance of global aquaculture production  

    a. Brief summary of ecological, social and economic benefits of aquaculture  
    
    b. Projected increase in seafood demand and aquaculture as the most likely means to fill this demand (i.e., limited ability of wild fisheries to increase production) 
    
    c. Despite strong arguments for aquaculture growth to meet increasing seafood demand at a global scale, how aquaculture development and growth can and should be realized at a regional scale requires an understanding of the sustainable and economically viable production potential of a region 
    
    d.	Description of factors that must be considered when determining aquaculture production potential of a given region ( see McKindsey et al. 2006; Dunn et al. 2013). 
      * physical
      * ecological
      * social
      * economic
  
II)	The Caribbean as a region where aquaculture development could be particularly beneficial   
    
    a. Caribbean seafood consumption and production (include information on seafood exports and imports)  
    
    b. Describe current aquaculture production occurring in the Caribbean (minimal)  
    
    c. Recently, further development of aquaculture has been identified as a priority in the Caribbean to reduce reliance on imports and provide a new source of income and livelihoods.  
    
    d. Potential for land-based and coastal aquaculture in the Caribbean is extremely limited (explain why…..limited space on land, limited freshwater, limited energy resources, environmental concerns with coastal aquaculture), suggesting the need to look to offshore aquaculture  
  
III) Offshore aquaculture is an emerging approach to mariculture where farms are some distance offshore  

    a. Overview of offshore aquaculture worldwide  
    
    b. Brief discussion of advantages of offshore aquaculture (specifically in the Caribbean)  
    
    c. Likely species/technology in the Caribbean ((concluding that cobia is one of the more promising species because it is fast growing, has had demonstrated success in offshore farms, is a high value species, etc. and thus is the focus of our analysis))  
    
IV) IV)	Overview of cobia biology and cobia farming (this section should make it clear why SST was not a necessary layer in our suitability assessment , but was important to consider in our growth model).
  
V) Introducing our study  

    a. Offshore aquaculture is nearly untapped in the Caribbean, raising questions about what the actual potential is, how that potential varies across the many EEZs of the Caribbean, and what factors are currently most likely to be limiting development 
    
    b. Using high resolution spatial data, we develop a spatial bioeconomic model to estimate the total production potential (in terms of weight and profit) of offshore aquaculture in the Caribbean, using cobia as a case study. We examine the spatial distribution of offshore aquaculture potential to identify ‘hotspots’ for potential future offshore aquaculture development, and how it is distributed across space. Additionally, we examine the variation of individual farm profitability between and across EEZs to determine the importance of site selection in aquaculture development and to quantify the potential benefits of strategic spatial planning.  We explore these results under several regionally scaled production capacity scenarios, and identify the main parameters constraining production in our model. Our approach could be applied to other farmed species or applied to other regions, and our results can help to chart a course for a sustainable and economically prosperous offshore aquaculture industry in the Caribbean. 


# Methods  

## Description of study region

The area of focus for this tudy is the Territorial waters and Exclusive Economic Zones (EEZs) surrounding the 28 island countries that comprise the Greater and Lesser Antilles of the Caribbean Sea (Figure 1). We did not consider the poteintial of offshore aquaculture development in the high seas for this stud. A shape file was obtained from VLIZ [-@vliz_maritime_2014] was used to define the maritime boundaries for all island countries. 

```{r Caribbean EEZ tmap, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="**Figure 1.** Th study area for this project is indicated in blue. EEZ boundary data were obtained from VLIZ (2014).",fig.width=6.5, fig.height=4}

library(tmap)
library(rgdal)
data(World)

land = readOGR( dsn="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/EEZ",layer="carib_eez_shape")
country<-land@data$GeoName[c(1,2,3,4,5,8,9,10,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,34,35,36,37,40,41,42,43,44,45,48)]
land$la<-ifelse(land@data$GeoName %in% country,"lightblue","white")
 tm_shape(land)+
  tm_fill("la")+
  tm_borders()+
   tm_legend()
```


## Suitability Assessment

The first step in our analysis was to conduct an assessment to identify areas that are potentially suitable for the development of offshore  mariculture. To determine what areas are suitable, we considered 6 factors: depth, current speed, shipping activity, oil sturctures, and areas designated for conservation (Table 1). 

We conducted our suitability assessment assuming the use of SeaStation cages, which are submersible, self-tensioned, single rim cages and are the most widely used offshore aquaculture cages in the U.S. [@loverich_case_2010; Open Blue 2016).The minimum site depth for installation of a SeaStation cage, listed by the original manufacturer, is 25 m (OceanSpar 2013). The maximum suitable depth for cage installation for this analysis was set at 90 m. In waters deeper than 90 m, cage installation and inspection of mooring and anchoring systems is more difficult and costly [@scott_offshore_2000]. It is possible in the future, deeper depths could become more feasible from an economic and engineering perspective, and this maximum depth limit could be adjusted accordingly in subsequent analyses. Spatial bathymetry data were obtained from Becker *et al* [-@becker_global_2009a], and areas that fell within a 25-90 m depth were scored as suitable for offshore mariculture development. Additionally, SeaStation cages can withstand consistent current velocities up to 1 m/s (Loverich 2007). To identify areas with unsuitably high current velocities, we obtained 10 years (2005-2015) of monthly average current data from Ocean Surface Current Analysis Real-time (OSCAR) zonal and meridional surface current velocity estimates from JPL Physical Oceanography DAAC and developed by Earth Space Research [@esr_oscar_2009]. The absolute maximum zonal and meridional velocities for each cell in our study area were calculated, and any cell with a maximum zonal or meridional velocity greater than 1 m/s were scored as unsuitable for the development of offshore mariculture.

To identify areas currently designated or utilized for other purposes, and for that reason unsuitable for offshore mariculture development, we obtained spatial data on: shipping activity, existing benthic oil structures, areas permitted for deep sea bed mining, and areas that have been designated as Marine Protected Areas or Conservation Priority areas (Table 1). Areas that fell in the 10% of the highest relative shipping activity, designated for conservation or deep sea bed mining purposes, and with existing benthic oil structures were considered unsuitable for offshore aquaculture development. 


All spatial data layers listed in Table 1 were converted to raster format with a 1 km^2^ spatial resolution. Data files that had a resolution > 1 km^2^ were interpolated to the 1 km^2^ grid using 'nearest neighbor technique' in R's raster package [@Hijmans_introduction_2012]. All 1 km^2 cells from each layer that fell outside of the suitable threshold, were given a score of zero. We then stacked all data layers in a boolean overlay to identify areas suitable for offshore mariculture development based on the criteria in Table 1.   


```{r kable, echo=FALSE, warning=FALSE, message=FALSE}
data<-read.csv("table_1.csv",check.names=FALSE)
#knitr::kable(head(data),caption="Table 1. Spatial data layers used in suitablility analysis to #identify cells suitable for cobia offshore mariculture",row.names=FALSE, #align=c('c','c','l','c'),padding=10)
library(pander)
set.alignment("center")
names(data) <- pandoc.strong.return(names(data))
pander::pandoc.table(data,justify=c('center','center','center','left','left','center'),split.cells=c(9,9,9,40,10,10),split.tables=Inf,caption=" Spatial data layers used in suitablility analysis to #identify cells suitable for cobia offshore mariculture")
```

## Farm Design

To estimate production, we assumed a fixed farm design across our study region. SeaStation cages are typically configured using a grid mooring system that includes grid, anchor, and mooring lines secured at varying distances from the cages [@fredriksson_design_2004; @xu_research_2015).We developed a hypothetical farm design for a 1 km^2^ cell that  has 16 SeaStation cages (each 6,400 m^3^) SeaStation cages configured in two eight-cell grid mooring systems that occupy a total space of approximately (~0.48 km^2^) and provide a total cage capacity of 102,400 m^3^ (Figure 2). The cages are held in position by the mooring system at depths of 15-20 m below the surface[@loverich_case_2010].

This cage configuration is similar to a 0.4 km^2^ offshore mariculture farm located off the coast of Kona, Hawaii that has an eight-cage SeaStation array with a total cage volume of 64,000 m^3^ (Sims et al. 2013). The total cage volume per unit of total farm area of this hypothetical design falls within the range of total cage volume per unit area of farm area for the Kona farm [@sims_kona_] and another offshore farm using SeaStation cages in the Gulf of Maine [@fredriksson_design_2004]; @decew_assessment_2010). Additionally, the total space occupied by the farm's infrastructure (0.48 km^2^) follows the guidelines issued in NOAA's Fishery Management Plan (FMP) for offshore aquaculture development in the Gulf of Mexico specifying
that total farm area should be twice the size of the total area occupied by farm infrastructure (GMFMC and NOAA 2009).

To maintain a steady supply of product, commercial aquaculture farms typically stage the stocking of cages so that indivdiuals of harvestable size are available year round. Cobia take on average, 12 months to grow from fingerling to a harvestable size (). We assume that the stocking of individual cages within a farm is staged at monthly time step. This means that for any given month, there will be between 1 and 5 cages with cobia at a harvestable age. We set the intial stock density of cages at 7.2 grams per m^3, which is equivalent to 2.3 juveniles with an average weight of 3 grams [@benetti_growth_2010].This stocking density has been shown to produce higher growth rates for cobia in the Caribbean than cages stocked at a higer density [@bennetti_growth_2010].

## Bioeconomic Model

We applied a bioeconomic model to estimate the production (in terms of cobia biomass yield (mt) and profit ($)) of individual farms developed in each of the 1 km^2^ cells that were identified as suitable for the development of offshore aquaculture.  Only farms that generated positive profit over a ten year period were ultimately considered feasible for development.

### Biological Model 
We used a thermal performance curve (TPC) to model temperature dependent growth of cobia (Klinger et al. ) Temperature is one of the primary abiotic factors controlling growth in fish, including cobia [@benetti_growth_2010], and is a factor that cannot be controlled in offshore mariculture grow-out settings. We used 10 years of satellite-based sea surface temperature (SST) data (NASA 2014) to calculate 1 km^2^ resolution spatial data layers of average monthly SST. 

We assume a t_min_ of 22 C and a t_max_ of 32 C (REF). Monthly biomass $BM$ is then calculated for each month, at each individual farm as:

* INSERT DANE'S THERMAL PERFORMANCE CURVE EQUATION *

where $D$, is the initial stocking density ($\frac{kg}{m^3}$ ) of each farm, and $M$, the monthly natural mortality rate is assumed fixed at XX.
 
Monthly biomass is then summed, to determine total annual biomass ($B_t$) at each farm.

Where $B_t$ estimates biomass (mt) in month $t$ as a function of average monthly sea surface temperature, assuming an initial stocking density of XX, and a natural mortality ($M$) rate of XX [@benetti_growth_2010]. 

###Economic Model
Using cost parameters we derived from the literature for each EEZ (Table 2), and parameters that were fixed across all sites (Table 3), the economic model estimates total annual operating costs ($TC_t$) for year 1 ($t = 1$) as:  
$TC_{t = 1} = SC + OC_{t = 1}$

Where $SC$ represents capital expenditures (start up costs) as:  

$SC_{t=1} = E_{depth} + I_{distance,EEZ} + P_{EEZ}$

where $E$ represents materials and equipment costs as a function of depth, I represent installation costs as a function of the farm's distance from shore and, base installation cost depnding on EEZ, and $P$ represents the cost of an aquaculture permit or lease as a function of EEZ (Table 2). 

$OC_t$ represents annual operating costs at each year $t$:

$OC_t = (u_{wave_exposure} * w * d_{EEZ}) + (\frac{distance}{b} * w * d_{EEZ} * r) + (distance * b * g_{EEZ} * e * t) + (f_t +h_t)$

Where $u$ is the hours required at each site as a function of wave exposure, $w$ is the number of workers required to service a farm, and $d$ is the hourly wage per worker as a function of EEZ, $distance$ is distance from shore (km), $b$ is average boat speed to site, $r$ is the number of trips required per month for running the farm, $g$ is the cost of fuel as function of EEZ, $e$ is the average fuel efficiency of the vessel, $f$ is the cost of fingerlings in year $t$, and $h$ is the cost of feed in year $t$. The global market for cobia is currently very small, with just 8.5% of production (4,555 MT) exported in 2012. We assume that cobia aquaculture in the Caribbean will be an export commodity and thus increased production in the Caribbean will affect the global price of cobia $c_t$ as follows:

$c_{t+1} = c_t - 0.8c_{t}(\frac{\Delta_{x}}{x_{t-1}})$

Where $c_{t+1}$ is the global price of cobia (\$) in the following year and $\frac{\Delta_{x}}{x_{t-1}}$ is the percent change in Caribbean production from the previous year. The constant 0.8 represents the price elasticity of demand and suggests that for every 1% increase in price there is a 0.8% decrease in consumption.

Total annual revenue, $TR$, for each year $t$,  for each farm was calculated as :

$TR_t = ( B_t c_t) – (B_t c_t j_{EEZ})$ 

Where $j$ is the export tax as a function of EEZ. 

The total profit ($TP$) for each farm over $t$ years was calculated as:

$TP = \displaystyle\sum_t^{10}TR_t -\biggl[\displaystyle\sum_t^{10}SC_t + \displaystyle\sum_t^{10}OC_t\biggr]$


We calculated the total summed biomass (mt) over a ten year period, and the 10 year Net Present Value (NPV), and only farms that generate positive revenue are considered feasible for development. Feasible farms within each EEZ were then ranked according to: 1. yield, and 2.profit (highest to lowest). We then examined the range of estimated yields and profits generated by farms within an EEZ, and compared the highest and lowest revenue generating farms across EEZs.

## Production Capacity (regional ecological carrying capacity)

Determining the aquaculture production level that a given region (in this case, an EEZ) can sustain without resulting in negative impacts to the surrounding ecosystem and environment is difficult, and will depend on the specific local oceanographic conditions and ecosystem dynamics. Although models have been developed to simulate conditions of a local area and to generate impacts of given aquaculture production scenarios on the surrounding ecosystem [REF], currently, limited empirical evidence exists to estimate the ecological carrying capacity for aquaculture production at a regional scale due to lack of extensive offshore aquaculture development.
```{r GOM FMP,echo=FALSE,warning=FALSE,message=FALSE}
Totalarea<-35643
Totalproduction<-29000
Productionrate<-Totalproduction/Totalarea
```

Guidelines for aquaculture development in U.S. waters of the Gulf of Mexico listed in a recent Fishery Management Plan stated that all aquaculture farms should be a minimum of 3 km apart and set a precautionary production limit of `r prettyNum(Totalproduction, big.mark=",")` mt for a total area of `r prettyNum(Totalarea,big.mark=",")` km^2^ identified for aquaculture developement, or `r prettyNum(Productionrate,big.mark=",")` mt/ km^2^. This production limit was identified as precautionary and based on historical wild fisheries capture for the region, due to lack of better methods or information on the ecological capacity of the region. To determine how this specified limit on production per unit area compared to limits applied in other regions, we conducted a global assessment of production per uit of suitable area for all EEZs. We obtained total marine aquaculture production by country, which included finfish, invertebrate, and algae aquaculture production, for 2014 from FAO and calculated the total area in a suitable depth range (0-90 m) for aquaculture by EEZ using global 1 km^2^ resolution bathymetry data from Becker *et al* [-@becker_global_2009a]. Total marine aquaculture production included finfish, invertebrates, and algae, thus We did not apply a minimum depth limit, because some these species are farmed in shalower water than finfish cage culture can happen. Although this not allow for a direct comparison to production modelled in our study, it provides another benchmark for comparison.

We then investigate a range of potential regional production limit scenarios at the EEZ scale for the Caribbean:

1. No production limit- There is no regionally specified production limit
2. High production limit- We set the production limit by identifying the EEZ worldwide with highest production per unit area globally from our  production assessment and multiply that value by the total area identfied as suitable for offshore aquaculture development in each EEZ in the Caribbean.
3. Precautionary limit- The production limit for each EEZ is specified by multiplying the production per unit area value specified for the Gulf of Mexico (`rProductionrate,big.mark=",") by the total area identified as suitable for offshore aquaculture development in  each EEZ.

Under each scenario, individual farms were developed in order of highest annual revenue. Once a single farm is developed, all farms falling within 3 km in any direction from that farm are excluded (not developed), to decrease the chance of negative cumulative enviornmental impacts on the surrounding ecosystem.

We then examined and compared the following under each production capacity scenario:
•	Total production (in terms of profit and weight) for the entire study (Caribbean) region
•	Total production (in terms of profit and weight) for each EEZ
•	Range of individual farm production within EEZs (to the see potential gains of strategic spatial planning) and across EEZs (to determine the island countries with the greatest potential for offshore aquaculture development)

Finally, we determined the main factors contributing to variability of farm and EEZ level aquaculture production by performing a suite of sensitivity analyses, in an effort to identify the current major hurdles to the development of offshore aquaculture in the Caribbean. 

	
# Results

##Suitability Assessment
```{r Suitabiliby Analysis, echo=FALSE,warning=FALSE,message=FALSE}
library(raster)
library(rgdal)

#This creates a spatial layer of 1km2 cells that may be suitable for cobia aquaculture, and estimates the number of suitable cells in each EEZ and accounted for by each factor

##Assigns a zero to international waters and 1 to waters inside EEZ
raw_eez<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/carib_eez_ocean.tif")
#codes<-read.csv("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/country_codes.csv")
non_countries<-c(54,55,62,63,98,108,109,125,126,128,271,272,273,274,108,164,129,130,131,132,133,135,139,163,167,255,221) 

for (i in 1:length(non_countries)){
 raw_eez[raw_eez==non_countries[i]]<-0
}
raw_eez[raw_eez==0]<-NA
eez<-raw_eez
eez[eez>0]<-1

###Assigns depth values to either a 1 (between 25-90 m depth) or 0
carib_depth<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/carib_depth.tif")
carib_depth<-mask(carib_depth,eez,inverse=FALSE,maskvalues=1)
carib_depth[is.na(carib_depth)]<-0
carib_depth[carib_depth<25]<-0
carib_depth[carib_depth>90]<-0
land<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/carib_depth.tif")
carib_depth<-mask(carib_depth,land,inverse=FALSE,maskvalues=NA,filename="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/suitable_depth.tif",overwrite=TRUE)
carib_depth[carib_depth!=0]<-1

###Assigns cells with MPAs present a zero
MPA<-raster('C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/final_mpa_layer.tif')
MPA<-mask(MPA,eez,inverse=FALSE,maskvalues=1)
land<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/carib_depth.tif")
suitableMPA<-mask(MPA,land,maskvalues=NA,inverse=FALSE,filename="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/suitable_MPA.tif",overwrite=TRUE)

###Assigns cells either a 1 (< 1 m/s )  or 0 based on current velocity
u<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/final_max_int_u.tif")
v<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/final_max_int_v.tif")
s<-stack(u,v)
current<-calc(s,fun=function(x){max(abs(x),na.rm=FALSE)})
current<-mask(current,eez,inverse=FALSE,maskvalues=1)
current<-mask(current,land,maskvalues=NA,inverse=FALSE,filename="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/final_carib_current.tif",overwrite=TRUE)
current[current>1]<-0
current[current==1]<-0
current[current!=0]<-1

###ssigns a 0 to cells with the top 10% of highest shipping activity
ships<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/carib_ship.tif")
ships[is.na(ships)]<-0.1
ships<-mask(ships,eez,inverse=FALSE,maskvalues=1)
top<-unique(ships) #5949970
top<-sort(top,decreasing=TRUE)
num<-(length(top)*.10)
top_num<-top[1:num]
ship_threshold<-min(top_num)
ships[ships>ship_threshold]<-0 ## 0.4 was chosen as an arbitary threshold for now
ships[ships==ship_threshold]<-0
ships<-mask(ships,land,maskvalues=NA,inverse=FALSE,filename="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/suitable_shipping.tif",overwrite=TRUE)
ships[ships>0]<-1

###Assigns a 0 to cells with the benthic oil structures
oil<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/suitable_oil.tif")
oil<-mask(oil,eez,inverse=FALSE,maskvalues=1)
oil[oil==1]<-100
oil[oil==0]<-1
oil[oil==100]<-0


##Create dataframe summarizing the total suitable area based on each factor (area function needs NAs)

factors<-c("EEZ","Depth","Maximum current velocity","Marine Protected Areas","Shipping Activity","Benthic Oil Structure","Deep Sea Mining")
df<-data.frame(matrix(nrow=7, ncol=3))
colnames(df)<-c("Factor", "Suitable Area (km2)", "Percent of Total Study Area")
df$Factor<-factors

###Total EEZ area
eez2<-eez
eez2[eez2==0]<-NA
eez_area<-as.data.frame(area(eez2,na.rm=TRUE))
eez_total_area<-sum(eez_area,na.rm=TRUE)
df[1,2]<-eez_total_area
df[1,3]<-100

###Total depth area
carib_depth2<-carib_depth
carib_depth2[carib_depth2==0]<-NA
depth_area<-as.data.frame(area(carib_depth2,na.rm=TRUE))
total_depth_area<-sum(depth_area,na.rm=TRUE)
df[2,2]<-total_depth_area
df[2,3]<-total_depth_area/eez_total_area*100

### Total MPA area
MPA2<-suitableMPA
MPA2[MPA2==0]<-NA
MPA_area<-as.data.frame(area(MPA2,na.rm=TRUE))
total_MPA_area<-sum(MPA_area,na.rm=TRUE)
df[4,2]<-total_MPA_area
df[4,3]<-total_MPA_area/eez_total_area*100

### Total current area
current2<-current
current2[current2==0]<-NA
current_area<-as.data.frame(area(current2,na.rm=TRUE))
total_current_area<-sum(current_area,na.rm=TRUE)
df[3,2]<-total_current_area
df[3,3]<-total_current_area/eez_total_area*100

### Total shipping area
ships2<-ships
ships2[ships2==0]<-NA
ship_area<-as.data.frame(area(ships2,na.rm=TRUE))
total_ship_area<-sum(ship_area,na.rm=TRUE)
df[5,2]<-total_ship_area
df[5,3]<-total_ship_area/eez_total_area*100

### Total oil area
oil2<-oil
oil2[oil2==0]<-NA
oil_area<-as.data.frame(area(oil2,na.rm=TRUE))
total_oil_area<-sum(oil_area,na.rm=TRUE)
df[6,2]<-total_oil_area
df[6,3]<-total_oil_area/eez_total_area*100

### Total Deep sea area
df[7,2]<-eez_total_area
df[7,3]<-eez_total_area/eez_total_area*100

###final dataframe
df<-df[order(df$`Suitable Area`,decreasing=TRUE),]


#Boolean Overlay to identify final suitable cells using all layers and calculate
s = stack(carib_depth,suitableMPA,current,ships,eez,oil)
all<-stackApply(s,indices=c(1,1,1,1,1,1),fun=prod,na.rm=FALSE)
all<-mask(all,land,maskvalues=NA,inverse=FALSE,filename="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/Suitable areas.tif",overwrite=TRUE)
all[all==0]<-NA
suitable_area<-as.data.frame(area(all,na.rm=TRUE))
final_suitable_area<-sum(suitable_area,na.rm=TRUE)
percent_total<-final_suitable_area/eez_total_area*100


# Calculate total suitable area by country
EEZ = readOGR(dsn="C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/EEZ",layer="carib_eez_shape2")
EEZ$PolygonID<-as.numeric(as.character(EEZ$PolygonID))
for (i in 1:length(non_countries)){
 EEZ$PolygonID[EEZ$PolygonID==non_countries[i]]<-0
}
EEZ$PolygonID[EEZ$PolygonID==0]<-NA

countries<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/tmp/carib_eez_ocean.tif")
for (i in 1:length(non_countries)){
 countries[countries==non_countries[i]]<-0
}
countries[countries==0]<-NA

#country<-raster("C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/suitable_eez.tif")
names<-data.frame(matrix(NA,nrow=38,ncol=3))

EEZ2<-as.data.frame(EEZ)
c<-EEZ2[,c(1,3,22)]
c<-na.omit(c)
c<-c[-38,]
area_all<-area(all,na.rm=TRUE)
eez_area<-zonal(area_all,countries,fun='sum')

i=1
for (i in 1:nrow(eez_area)){
m<-as.integer(eez_area[i,1])
names[i,1]<-m
n<-c$GeoName[c$PolygonID==m]
n<-as.character(n[1])
names[i,2]<-n
a<-c$Area_km2[c$PolygonID==m]
a<-as.integer(a[1])
names[i,3]<-a
}

cells<-cbind(eez_area,names,c$Area_km2)
cells<-cells[,-3]
cells[order(cells$sum,decreasing=TRUE),]
write.csv(cells,"C:/Users/Lennon Thomas/Desktop/Carib_aqua_16/Suitability/final/suitable area by country.csv")

####Plot suitable area using tmaps######
test<-select(country)
library(tmap)
data(World)

land = readOGR(dsn="tmp",layer="carib_eez_shape")
land$la<-ifelse(land@data$rgn_type=="land","pink","white")

suitable<-tm_shape(land)+
  tm_fill("la")+
tm_shape(all)+
  tm_raster(showNA=FALSE,legend.show=FALSE)+
tm_shape(land)+
  tm_borders()

save_tmap(suitable,filename="datalayer plots/suitability.png")
tm.layout(legend.position=c("right","top"))


```

Based on the factors and criteria described in Table 1, we found a total of XX km^2^ of area suitable for aquaculture development in the Caribbean (Figure 2).  The total area available based on the criteria for each factor are listed in Table 4. XX was the largest constraining factor in detemrining areas suitable for offshore aquaculture. Deep sea mining did not eliminate any areas because currently no deep sea bed mining permits have been issued in the region. However, this may be an important factor to consider in other regions. 

The identified suitable areas were not distributed evenly across EEZs (Table 5). XX included XX% of the total suitable area, and XX EEZs did not contain any areas that were identified as suitable.

##Production Capacity

##Production capacity
### Regional
### By EEZ

# Discussion
#References

NASA Goddard Space Flight Center, Ocean Ecology Laboratory, Ocean Biology Processing Group; (2014): MODIS-Terra Daily 4 km Sea Surface Temperature Data; http://doi.org/{DOI} Accessed on January 3, 2016.

GMFMC (Gulf of Mexico Fishery Management Council)and NOAA (National Oceanic and Atmospheric Administration)(2009) Fishery management plan for regulating
offshore marine aquaculture in the Gulf of Mexico. GMFMC, Tampa, FL.http://gulfcouncil.org/Beta/GMFMCWeb/Aquaculture/Aquaculture%20FMP%20PEIS%20Final%202-24-09.pdf

OceanSpar 2013.